---
title: "Data Envelopment Analysis (DEA)"
author: "Muthia Salsabila"
output:
  html_document: 
    toc: yes
---
<br></br>
This Notebook show how to estimate and measure efficiency using Data Envelopment Analysis (DEA) method and Benchmarking package in R. 

         
<font size="5">What is Efficiency?</font>        
Efficiency is when a firm can minimize its input to produce a particular output. There are three types of efficiency, which are technical efficiency, allocative efficiency, and economic efficiency. Technical efficiency is a firm ability to reach optimal output using a particular input. Allocative efficiency is a firm ability to use optimal input to reach a certain level of output. Economic efficiency is a combination between technical efficiency and allocative efficiency.

Efficiency can be measure with DEA and SFA. We will be using DEA method for measuring the efficiency.

<font size="5">What is Data Envelopment Analysis?</font>     
Data Envelopment Analysis (DEA) is a non-parametric linear programming approach for measuring efficiency that can measure multiple-input and multiple-output. DEA introduced by Charnes, Coopers, and Rhodes in 1978. 

There are two assumptions in DEA, which are Constant Return to Scale (CRS) and Variable Return to Scale (VRS).    
<ul>
<li> CRS assume a firm operate at optimum scale and the ratio of increasing input is equal to output, which mean if there is an addition to input x times then the output will increase x times too. CRS model often define as CCR model because it introduced by Charnes, Cooper, and Rhodes.</li>           
<li> VRS (or known as BCC model, developed by Banker, Charnes, and Cooper) assume a firm not or not yet operate at optimum scale and the ratio between input and ouput is not equal. When there is an addition to input x times, the output can increase more than x or lower than x.</li>
</ul>

There's also two approach in DEA:
<ul>
<li>Input oriented = minimizing input to reach a certain level of output</li>
<li>output oriented = maximizing output using a particular combination of input</li>
</ul>

<font size="5">Step of Project</font>      
1. Loading Packages    
2. Import Data    
3. Select Input and Output    
4. Estimate Efficiencies    
5. Measure Slack    
6. Plot Frontier    

<font size="5">Loading Packages</font>     
The first step is to load the Benchmarking package using library function. This package contains methods to estimate and measure efficiency using DEA and SFA. 
```{r, message=FALSE, warning=FALSE}
library(Benchmarking)
```

<font size="5">Import data</font>    
This package also contains a few datasets that can be used for practice. The datasets I use called *milkProd*, which is a data from Danish milk producers. I also use dplyr package to see a glimpse of data rows, column, and type.
```{r}
#import data
data("milkProd")
milkProd
```
```{r}
dplyr::glimpse(milkProd)
```
As we can see, the data has 108 rows and 5 columns with all variables data type are integer. The variable name meaning are as follows:     

- **farmNO** =  farm number     
- **milk** = Output of milk, kg     
- **energy** = Energy Expenses     
- **vet** = Veterinary expenses     
- **cows** = Number of Cows     

<font size="5">Select Input and Output</font>     
To estimate the efficiency, we must select which variables from the data are input and which are the output. DEA can select multiple output and multiple input, but in this case we use only one output and multiple input.

```{r}
# input output selection
x <- with(milkProd, cbind(energy, vet, cows))
y <- with(milkProd, cbind(milk))
```
X define the input variables which are energy expenses, veterinary expenses, and number of cows. While y define the output variable which is milk in kilogram.

<font size="5">Calculate Efficiency</font>     
After we select the input and output, we can calculate the efficiency using dea function as follows:
```{r}
#calculating efficiency
#VRS (bcc)
bcc <- dea(x, y, RTS = "vrs", ORIENTATION = "in")
eff(bcc)
```
We use BCC model and input-oriented.
```{r}
summary(bcc)
```
From the summary we can see there are 16 DMUs categorized as efficient, and the mean efficiency is 0.868.

We also can see how much DMUs become a peers for other DMU. As we see below, L34 or DMU number 34 become peers for 80 other DMUs.
```{r}
get.number.peers(bcc)
```
Using get.which.peers function, we can see DMUs group by its peer.
```{r}
get.which.peers(bcc)
```
While below is a list of DMU with their reference peer. DMU number 2 follow DMU 29, DMU 34, and DMU 105 as an example so DMU number 2 can be efficient.
```{r}
print(peers(bcc,NAMES=TRUE),quote=FALSE)
```

<font size = "5"> Measure Slack </font>        
By measuring slack, we can know which DMU need to reduce its input to achieve efficiency.
If the result show "FALSE", it means the DMU does not need to adjust the input. But if the result show "TRUE", it means the DMU need to adjust its input by decreasing it to achieve efficiency.

```{r}
s1 <- slack(x, y, bcc)
t <- data.frame(milkProd$farmNo, eff(bcc), s1$slack, s1$sx)
t
```
As we can see the DMU that already achieve efficiency does not need to adjust its input.

Let's take a look at other DMU that still not efficient. DMU 2 need to reduce veterinary expenses by 14137.8443 to achieve efficiency. Meanwhile, DMU 3 need to reduce energy expense by 45551.15902 and veterinary expense by 29773.6018 to achieve efficiency.
If we take a deep look at the list, we can see there's not much result that show we should reduce cow input. It is because reducing the number of cow not making sense. But, we can reduce the expenses of energy and veterinary to minimize the input and achieve optimal output.

<font size = "5">Plot Frontier</font>      
We can visualize the result by plotting the DMU in basic plot of frontier. The red line is frontier for Variable return to scale, and the green dash line is constant return to scale.
```{r}
dea.plot(x,y, txt=1:dim(x)[1],main="Basic plot of frontier", col = "red")
dea.plot(x,y,RTS="crs",add=TRUE,lty="longdash", lwd="2", col="green")
```

<font size = "5">Conclusion</font>           
The number of Farm that produce milk efficiently are still lower than the inefficient one. There are only 16 farms that produce milk efficiently, while other 92 farms still produce milk inefficiently. To achieve efficiency, farm can reduce its input which are energy expenses and veterinary expenses but there is no need to reduce the number of cows because it does not make sense to reduce the cows unless they die. 